Optimal design of experiment: supercritical fluid extraction case

### 2024-08-04 SFE_poster_DOE/conference_poster_6.tex
04.08

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Optimal design of experiment:} \color{Black}\\ % Title
\Huge\textit{supercritical fluid extraction case}\\[2cm] % Subtitle
\huge \textbf{Oliwer Sliczniuk \& Pekka Oinas}\\[0.5cm] % Author(s)
\huge School of Chemical Engineering, Aalto University, Finland\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=15cm]{Aalto_University_logo.png}\\ \\
\end{minipage}

@@

This study investigates the extraction of valuable components from biomass using supercritical carbon dioxide as a solvent. The interest is in essential oils extracted from chamomile flowers in cylindrical extractors operated in a semi-batch mode. The extraction process is described by a distributed-parameter model, and set of empirical correlations. This work aims \underline{to validate the empirical correlations by designing a new experiment} with dynamically changing operating conditions. %The process model includes empirical correlations, which describe the relation between the extraction kinetics parameters and the operating conditions. 
\end{tcolorbox}

@@

%\vspace{0.1cm}
%\section*{Process Model}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Process Model\\}

\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item One-dimensional
		\item Plug flow
		\item No pressure drop
		\item Uniform particle distribution
	\end{itemize}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item Based on the two-film theory for a single component
		\item Peng-Robinson equation of state
		\item Decaying extraction kinetic
		\item Empirical correlations
	\end{itemize}
\end{minipage}

@@

\begin{align*}
	\dot{x} = \cfrac{d x}{d t} = 
	\begin{bmatrix}
		\frac{\partial c_f}{\partial t} \\
		\\
		\cfrac{\partial c_s}{\partial t} \\
		\\
		\cfrac{\partial \left(\rho_f h A_f\right)}{\partial t} \\
		\\
		\cfrac{d y}{d t}
	\end{bmatrix}
	=
\underbrace{\begin{bmatrix}
		-\frac{1}{\phi} \frac{\partial \left( c_f u\right)}{\partial z} + \frac{1-\phi}{\phi} \cfrac{\partial c_s}{\partial t} + \frac{1}{\phi} \frac{\partial}{\partial z} \left( D^M_e \frac{\partial c_f}{\partial z} \right) \\
		-\cfrac{ D_i^R \exp \left( \Upsilon \left( 1-\cfrac{ c_s }{c_{s0}} \right) \right) }{ \mu l^2 } \left( c_s  - \cfrac{\rho_s c_f }{ k_m \rho_f }  \right) \\
		- \cfrac{\partial \left( \rho_f h A_f {\color{red}v} \right)}{\partial z} + \cfrac{\partial \left(P A_f\right)}{\partial t} + \cfrac{\partial}{\partial z} \left( k \cfrac{\partial T}{\partial z} \right) \\
		\cfrac{F}{\rho_f} c_f \biggr\rvert_{z=L}
	\end{bmatrix}}_{G(x,t,\Theta;\Xi)}
\end{align*}

\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		c_f &- \text{Soluts concentration in the fluid phase}\\
		c_s &- \text{Soluts concentration in the solid phase}\\
		h~  &- \text{Enthalpy}\\
		y~ &- \text{Extraction yield}\\
		\rho_f &- \text{Density of fluid}\\
		A_f &- \text{Cross-section of the bed}\\
		u~~ &- \text{Darcy velocity}\\
		\phi~~ &- \text{Void fraction}
	\end{align*}
\end{minipage}
%
\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		D_e^M &- \text{Axial mass diffusivity}\\
		D_i^R &- \text{Internal mass diffusivity}\\
		\Upsilon~~ &- \text{Decaying factor}\\
		\rho_s~ &- \text{Bulky density of solid bed}\\
		~k_m &- \text{Partition factor}\\
		P~~ &- \text{Pressure}\\
		T~~ &- \text{Temperature}\\
		F~~ &- \text{Mass flow rate}
	\end{align*}
\end{minipage}

\begin{center}%\vspace{1cm}
	\includegraphics[width=0.49\linewidth]{Di_Re_F_1.png}
	\includegraphics[width=0.49\linewidth]{Gamma_Re_F_1.png}
	\captionof{figure}{Empirical correlations}
\end{center}%\vspace{1cm}

@@

\underline{Fisher information $\mathcal{F}$ measures the amount of information} that an observable random variable carries about an unknown parameter of a distribution that models the random variable. $\mathcal{F}$ is defined as Hessian of the log-likelihood function with respect to the parameters $\Theta$, given the covariance matrix $\Sigma$. %Based on the Cramer-Rao inequality, $\mathcal{F}$ can used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

The optimal design of experiments is a concept that refers to planning an experiment, which allow parameters to be estimated without bias and with minimum variance. The D-optimality criterion is selected as the objective function, which leads to \underline{the minimisation of the volume of the ellipsoidal confidence region of parameter estimates} given the experimental condition $\Xi$.

@@

	\begin{center}%\vspace{1cm}
		\includegraphics[width=\linewidth]{scatter.png}
		\captionof{figure}{Values of the objective function}
	\end{center}%\vspace{1cm}

@@

	\begin{center}%\vspace{1cm}
		\includegraphics[width=0.49\linewidth]{Profile_T.png}
		\includegraphics[width=0.49\linewidth]{Profile_F.png}
		\captionof{figure}{Optimal profiles}
	\end{center}%\vspace{1cm}

@@

\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out

\section*{Conclusions}

\begin{itemize}
\item Pellentesque eget orci eros. Fusce ultricies, tellus et pellentesque fringilla, ante massa luctus libero, quis tristique purus urna nec nibh. Phasellus fermentum rutrum elementum. Nam quis justo lectus.
\item Vestibulum sem ante, hendrerit a gravida ac, blandit quis magna.
\item Donec sem metus, facilisis at condimentum eget, vehicula ut massa. Morbi consequat, diam sed convallis tincidunt, arcu nunc.
\item Nunc at convallis urna. isus ante. Pellentesque condimentum dui. Etiam sagittis purus non tellus tempor volutpat. Donec et dui non massa tristique adipiscing.
\end{itemize}

\color{DarkSlateGray} % Set the color back to DarkSlateGray for the rest of the content

>>>>

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Optimal design of experiment:} \color{Black}\\ % Title
\Huge\textit{supercritical fluid extraction case}\\[2cm] % Subtitle
\huge \textbf{Oliwer Sliczniuk, Pekka Oinas}\\[0.5cm] % Author(s)
\huge School of Chemical Engineering, Aalto University, Finland\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=15cm]{Aalto_University_logo.png}\\ \\
\end{minipage}

@@

This study investigates the extraction of valuable components from biomass using supercritical carbon dioxide as a solvent. The interest is in essential oils extracted from chamomile flowers in cylindrical extractors operated in a semi-batch mode. The process is described by a distributed-parameter model, and set of empirical correlations. This work aims \underline{to validate empirical} \underline{correlations by designing a new experiment with dynamically changing operating conditions}. %The process model includes empirical correlations, which describe the relation between the extraction kinetics parameters and the operating conditions. 
\end{tcolorbox}

@@

%\section*{Process Model}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Process Model\\}

\vspace{0.5cm}

\begin{minipage}[b]{0.5\linewidth}
	\begin{itemize}
		\item One-dimensional
		\item Plug flow
		\item No pressure drop
		\item Uniform particle distribution
	\end{itemize}
\end{minipage}
\hfill
\begin{minipage}[b]{0.5\linewidth}
	\begin{itemize}
		\item Two-film theory for a single component
		\item Peng-Robinson equation of state
		\item Decaying extraction kinetic
		\item Empirical correlations
	\end{itemize}
\end{minipage}

@@

\begin{align*}
	\dot{x} = \cfrac{d x}{d t} = 
	\begin{bmatrix}
		\frac{\partial c_f}{\partial t} \\
		\\
		\cfrac{\partial c_s}{\partial t} \\
		\\
		\cfrac{\partial \left(\rho_f h A_f\right)}{\partial t} \\
		\\
		\cfrac{d y}{d t}
	\end{bmatrix}
	=
\underbrace{\begin{bmatrix}
		-\frac{1}{\phi} \frac{\partial \left( c_f u\right)}{\partial z} - \frac{1-\phi}{\phi} \cfrac{\partial c_s}{\partial t} + \frac{1}{\phi} \frac{\partial}{\partial z} \left( D^M_e \frac{\partial c_f}{\partial z} \right) \\
		-\cfrac{ D_i^R \exp \left( \Upsilon \left( 1-\cfrac{ c_s }{c_{s0}} \right) \right) }{ \mu l^2 } \left( c_s  - \cfrac{\rho_s c_f }{ k_m \rho_f }  \right) \\
		- \cfrac{\partial \left( \rho_f h A_f {\color{red}v} \right)}{\partial z} + \cfrac{\partial \left(P A_f\right)}{\partial t} + \cfrac{\partial}{\partial z} \left( k \cfrac{\partial T}{\partial z} \right) \\
		\cfrac{F}{\rho_f} c_f \biggr\rvert_{z=L}
	\end{bmatrix}}_{G(x,t,\Theta;\Xi)}
\end{align*}

\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		c_f &- \text{Solutes concentration in the fluid phase}\\
		c_s &- \text{Solutes concentration in the solid phase}\\
		h~  &- \text{Enthalpy}\\
		y~ &- \text{Extraction yield}\\
		\rho_f &- \text{Density of fluid}\\
		A_f &- \text{Cross-section of the bed}\\
		u~~ &- \text{Darcy velocity}\\
		\phi~~ &- \text{Void fraction}\\
		D_i^R &- \text{Internal mass diffusivity}
	\end{align*}
\end{minipage}
%
\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		D_e^M &- \text{Axial mass diffusivity}\\
		\mu~ &- \text{Particle shape coefficient}\\
		l~ &- \text{Particle length}\\
		\Upsilon~ &- \text{Decaying factor}\\
		\rho_s &- \text{Bulk density of solid bed}\\
		k_m &- \text{Partition factor}\\
		P~ &- \text{Pressure}\\
		T~ &- \text{Temperature}\\
		F~ &- \text{Mass flow rate}
	\end{align*}
\end{minipage}

\begin{center}\vspace{0.5cm}
	\includegraphics[width=0.49\linewidth]{Di_Re_F_1.png}
	\includegraphics[width=0.49\linewidth]{Gamma_Re_F_1.png}
	\captionof{figure}{Empirical correlations}
\end{center}%\vspace{1cm}

@@

\underline{Fisher information $\mathcal{F}$ measures the amount of information} that an observable random variable carries about an unknown parameter of a distribution that models the random variable. $\mathcal{F}$ is defined as Hessian of the likelihood function with respect to parameters $\Theta$, given the covariance matrix $\Sigma$. %Based on the Cramer-Rao inequality, $\mathcal{F}$ can used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

The optimal design of experiments is a concept that refers to planning an experiment, which allow parameters to be estimated without bias and with minimum variance. The D-optimality criterion is selected as the objective function, which leads to \underline{the minimisation of the volume} \underline{of the ellipsoidal confidence region of parameter estimates} given the experimental condition $\Xi$.

@@

	\begin{center}\vspace{0.5cm}
		\includegraphics[width=\linewidth]{scatter.png}
		\captionof{figure}{Values of the objective function}
	\end{center}%\vspace{1cm}

@@

	\begin{center}\vspace{0.5cm}
		\includegraphics[width=\linewidth]{Profiles.png}
		\captionof{figure}{Optimal profiles}
	\end{center}%\vspace{1cm}

@@

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Conclusions\\}

	\begin{itemize}
		\item The optimal profiles of controls are similar for all the cases
		\item Low values of the objective are obtained at pressures close to supercritical point, where the variation in the inlet temperature cause large deviation in the physical properties of CO$_2$ and Reynolds number
		\item The mass flow rate is used as main control variable, which suggest that the system is more sensitive to the mass flow rate than the inlet temperature
	\end{itemize}
